---
title: "GO Slim Characterization"
author: "Yaamini Venkataraman"
date: "4/13/2020"
output: html_document
---

In this R Markdown document, I will characterize GO Slim terms associated with all detected genes (gene background) and differentially expressed genes (DEG).

# Set up R Markdown file

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Install packages

```{r}
#install.packages("dichromat")
require(dichromat)
```

# Obtain session information

```{r}
sessionInfo() #Obtain session information
```

# *Z. marina*

## Import data

```{r}
zosteraBGGOSlim <- read.table("Zostera_full_blastquery-GOslim-BP.sorted.unique.noOther", header = FALSE, sep = "\t") #GO Slim terms for gene background
colnames(zosteraBGGOSlim) <- c("geneID", "GOSlim") #Add column names
head(zosteraBGGOSlim)
```

```{r}
zosteraDEGGOSlim <- read.table("Zostera_blastquery-GOslim-BP.sorted.unique.noOther", header = FALSE, sep = "\t") #GO Slim terms for gene background
colnames(zosteraDEGGOSlim) <- c("geneID", "GOSlim") #Add column names
head(zosteraDEGGOSlim)
```

## Format data

```{r}
zosteraBGGOSlimCounts <- as.data.frame(table(zosteraBGGOSlim$GOSlim)) #Count the frequency of each GO Slim term
colnames(zosteraBGGOSlimCounts) <- c("BGGOSlim", "frequency") #Rename columns
head(zosteraBGGOSlimCounts) #Confirm counts
```

```{r}
zosteraBGGOSlimCounts$percentage <- (zosteraBGGOSlimCounts$frequency/sum(zosteraBGGOSlimCounts$frequency)*100) #Create a new column with percentages of GOSlim terms instead of raw counts
head(zosteraBGGOSlimCounts) #Confirm column creation
max(zosteraBGGOSlimCounts$percentage) #Obtain the max percentage to inform plot axes
```

```{r}
zosteraBGGOSlimCounts <- zosteraBGGOSlimCounts[c(4, 1, 5, 11, 13,
                                                 2, 6, 3,
                                                 12,
                                                 7, 10, 9, 8),] #Reorganize rows based on broader functions. Rows are general activity, development, stress response, and metabolic processes
head(zosteraBGGOSlimCounts) #Confirm organization
```

```{r}
zosteraBGGOSlimCounts$GOSlim <- c("Cell-Cell Signaling", "Cell Adhesion", "Death", "Signal Transduction", "Transport",
                                   "Cell Cycle and Proliferation", "Developmental Processes", "Cell Organization and Biogenesis",
                                   "Stress Response",
                                   "DNA Metabolism", "RNA Metabolism", "Protein Metabolism", "Other Metabolic Processes") #Capitalize labels
head(zosteraBGGOSlimCounts) #Confirm capitalization
```

```{r}
zosteraDEGGOSlimCounts <- as.data.frame(table(zosteraDEGGOSlim$GOSlim)) #Count the frequency of each GO Slim term
colnames(zosteraDEGGOSlimCounts) <- c("DEGGOSlim", "frequency") #Rename columns
head(zosteraDEGGOSlimCounts) #Confirm counts
```

```{r}
zosteraDEGGOSlimCounts$percentage <- (zosteraDEGGOSlimCounts$frequency/sum(zosteraDEGGOSlimCounts$frequency)*100) #Create a new column with percentages of GOSlim terms instead of raw counts
head(zosteraDEGGOSlimCounts) #Confirm column creation
max(zosteraDEGGOSlimCounts$percentage) #Obtain the max percentage to inform plot axes
```

```{r}
zosteraDEGGOSlimCounts <- zosteraDEGGOSlimCounts[c(4, 1, 5, 11, 13,
                                                 2, 6, 3,
                                                 12,
                                                 7, 10, 9, 8),] #Reorganize rows based on broader functions. Rows are general activity, development, stress response, and metabolic processes
head(zosteraDEGGOSlimCounts) #Confirm organization
```

```{r}
zosteraDEGGOSlimCounts$GOSlim <- c("Cell-Cell Signaling", "Cell Adhesion", "Death", "Signal Transduction", "Transport",
                                   "Cell Cycle and Proliferation", "Developmental Processes", "Cell Organization and Biogenesis",
                                   "Stress Response",
                                   "DNA Metabolism", "RNA Metabolism", "Protein Metabolism", "Other Metabolic Processes") #Capitalize labels
head(zosteraDEGGOSlimCounts) #Confirm capitalization
```

## Create color palette

```{r}
RColorBrewer::display.brewer.all() #Show all RColorBrewer palettes.
plotColors <- rev(RColorBrewer::brewer.pal(5, "YlGn")) #Create a color palette for the barplots. Use 5 yellow-green shades from RColorBrewer. Reverse the order so the darkest shade is used first.
barplot(zosteraBGGOSlimCounts$percentage,
        col = plotColors) #See what plot looks like with new scheme
barplot(zosteraBGGOSlimCounts$percentage,
        col = dichromat(plotColors)) #Check that the plot colors will be easy to interpret for those with color blindess
```

## Create figure

```{r}
#pdf("Zostera-BP-GOSlim-BG-Versus-DEG.pdf", width = 11, height = 8.5) #Save as a pdf

par(mfrow = c(1, 2), oma = c(5, 20, 0, 3), mar = c(0, 0, 0, 0)) #Set up parameters for multipanel plot

barsGOSlimBG <- barplot(-1*(zosteraBGGOSlimCounts$percentage),
                      horiz = TRUE,
                      axes = FALSE,
                      xlim = c(-20,0),
                      col = plotColors[4]) #Create a barplot that is horizontal (horiz = TRUE). Use axes = FALSE to remove all axes. Set xlim based on maximum percentage value. Colors correspond to broader GOSlim groupings. Save plot as a new object.

axis(side = 2, at = barsGOSlimBG, labels = zosteraBGGOSlimCounts$GOSlim, tick = FALSE, las = 2, col = "grey80", cex.axis = 1.5) #Add y-axis with GOSlim terms from BPGOSlimCounts. Place labels at specific barplot values from barsGOSlim. Remove tick marks (tick) and change orientation of labels (las)
axis(side = 1, at = seq(from = -20, to = 0, by = 5), labels = c(seq(from = 20, to = 0, by = -5)), cex = 1.2, col = "grey80") #Add x-axis
mtext(side = 1, "% BG", line = 3, cex = 1.5) #Add x-axis label

barsGOSlimDEG <- barplot(zosteraDEGGOSlimCounts$percentage,
                         horiz = TRUE,
                         axes = FALSE,
                         xlim = c(0,20),
                         col = plotColors[1]) #Create a barplot that is horizontal (horiz = TRUE). Use axes = FALSE to remove all axes. Set xlim based on maximum percentage value. Colors correspond to broader GOSlim groupings. Save plot as a new object.
axis(side = 1, at = seq(from = 0, to = 20, by = 5), cex = 1.2, col = "grey80") #Add x-axis
mtext(side = 1, "% DEG", line = 3, cex = 1.5) #Add x-axis label

#dev.off()
```

# *L. zosterae*

## Import data

```{r}
nonZosteraBGGOSlim <- read.table("nonZostera_full_blastquery-GOslim-BP.sorted.unique.noOther", header = FALSE, sep = "\t") #GO Slim terms for gene background
colnames(nonZosteraBGGOSlim) <- c("geneID", "GOSlim") #Add column names
head(nonZosteraBGGOSlim)
```

```{r}
nonZosteraDEGGOSlim <- read.table("nonZostera_blastquery-GOslim-BP.sorted.unique.noOther", header = FALSE, sep = "\t") #GO Slim terms for gene background
colnames(nonZosteraDEGGOSlim) <- c("geneID", "GOSlim") #Add column names
head(nonZosteraDEGGOSlim)
```

## Format data

```{r}
nonZosteraBGGOSlimCounts <- as.data.frame(table(nonZosteraBGGOSlim$GOSlim)) #Count the frequency of each GO Slim term
colnames(nonZosteraBGGOSlimCounts) <- c("BGGOSlim", "frequency") #Rename columns
head(nonZosteraBGGOSlimCounts) #Confirm counts
```

```{r}
nonZosteraBGGOSlimCounts$percentage <- (nonZosteraBGGOSlimCounts$frequency/sum(nonZosteraBGGOSlimCounts$frequency)*100) #Create a new column with percentages of GOSlim terms instead of raw counts
head(nonZosteraBGGOSlimCounts) #Confirm column creation
max(nonZosteraBGGOSlimCounts$percentage) #Obtain the max percentage to inform plot axes
```

```{r}
nonZosteraBGGOSlimCounts <- nonZosteraBGGOSlimCounts[c(4, 1, 5, 11, 13,
                                                 2, 6, 3,
                                                 12,
                                                 7, 10, 9, 8),] #Reorganize rows based on broader functions. Rows are general activity, development, stress response, and metabolic processes
head(nonZosteraBGGOSlimCounts) #Confirm organization
```

```{r}
nonZosteraBGGOSlimCounts$GOSlim <- c("Cell-Cell Signaling", "Cell Adhesion", "Death", "Signal Transduction", "Transport",
                                   "Cell Cycle and Proliferation", "Developmental Processes", "Cell Organization and Biogenesis",
                                   "Stress Response",
                                   "DNA Metabolism", "RNA Metabolism", "Protein Metabolism", "Other Metabolic Processes") #Capitalize labels
head(nonZosteraBGGOSlimCounts) #Confirm capitalization
```

```{r}
nonZosteraDEGGOSlimCounts <- as.data.frame(table(nonZosteraDEGGOSlim$GOSlim)) #Count the frequency of each GO Slim term
colnames(nonZosteraDEGGOSlimCounts) <- c("DEGGOSlim", "frequency") #Rename columns
tail(nonZosteraDEGGOSlimCounts) #Confirm counts
```

```{r}
nonZosteraDEGGOSlimCounts$percentage <- (nonZosteraDEGGOSlimCounts$frequency/sum(nonZosteraDEGGOSlimCounts$frequency)*100) #Create a new column with percentages of GOSlim terms instead of raw counts
head(nonZosteraDEGGOSlimCounts) #Confirm column creation
max(nonZosteraDEGGOSlimCounts$percentage) #Obtain the max percentage to inform plot axes
```

```{r}
nonZosteraDEGGOSlimCounts <- nonZosteraDEGGOSlimCounts[c(4, 1, 5, 11, 13,
                                                 2, 6, 3,
                                                 12,
                                                 7, 10, 9, 8),] #Reorganize rows based on broader functions. Rows are general activity, development, stress response, and metabolic processes
head(nonZosteraDEGGOSlimCounts) #Confirm organization. Row 13 is not present in this dataset
```

```{r}
nonZosteraDEGGOSlimCounts$percentage[5] <- 0 #Add a 0 to the % column
head(nonZosteraDEGGOSlimCounts) #Confirm addition
```

```{r}
nonZosteraDEGGOSlimCounts$GOSlim <- c("Cell-Cell Signaling", "Cell Adhesion", "Death", "Signal Transduction", "Transport",
                                   "Cell Cycle and Proliferation", "Developmental Processes", "Cell Organization and Biogenesis",
                                   "Stress Response",
                                   "DNA Metabolism", "RNA Metabolism", "Protein Metabolism", "Other Metabolic Processes") #Capitalize labels
head(nonZosteraDEGGOSlimCounts) #Confirm capitalization
```

## Create color palette

```{r}
RColorBrewer::display.brewer.all() #Show all RColorBrewer palettes.
plotColors <- rev(RColorBrewer::brewer.pal(5, "YlOrBr")) #Create a color palette for the barplots. Use 5 yellow-orange-brown shades from RColorBrewer. Reverse the order so the darkest shade is used first.
barplot(nonZosteraBGGOSlimCounts$percentage,
        col = plotColors) #See what plot looks like with new scheme
barplot(nonZosteraBGGOSlimCounts$percentage,
        col = dichromat(plotColors)) #Check that the plot colors will be easy to interpret for those with color blindess
```

## Create figure

```{r}
#pdf("nonZostera-BP-GOSlim-BG-Versus-DEG.pdf", width = 11, height = 8.5) #Save as a pdf

par(mfrow = c(1, 2), oma = c(5, 20, 0, 3), mar = c(0, 0, 0, 0)) #Set up parameters for multipanel plot

barsGOSlimBG <- barplot(-1*(nonZosteraBGGOSlimCounts$percentage),
                      horiz = TRUE,
                      axes = FALSE,
                      xlim = c(-21,0),
                      col = plotColors[4]) #Create a barplot that is horizontal (horiz = TRUE). Use axes = FALSE to remove all axes. Set xlim based on maximum percentage value. Colors correspond to broader GOSlim groupings. Save plot as a new object.

axis(side = 2, at = barsGOSlimBG, labels = nonZosteraBGGOSlimCounts$GOSlim, tick = FALSE, las = 2, col = "grey80", cex.axis = 1.5) #Add y-axis with GOSlim terms from BPGOSlimCounts. Place labels at specific barplot values from barsGOSlim. Remove tick marks (tick) and change orientation of labels (las)
axis(side = 1, at = seq(from = -21, to = 0, by = 7), labels = c(seq(from = 21, to = 0, by = -7)), cex = 1.2, col = "grey80") #Add x-axis
mtext(side = 1, "% BG", line = 3, cex = 1.5) #Add x-axis label

barsGOSlimDEG <- barplot(nonZosteraDEGGOSlimCounts$percentage,
                         horiz = TRUE,
                         axes = FALSE,
                         xlim = c(0,21),
                         col = plotColors[1]) #Create a barplot that is horizontal (horiz = TRUE). Use axes = FALSE to remove all axes. Set xlim based on maximum percentage value. Colors correspond to broader GOSlim groupings. Save plot as a new object.
axis(side = 1, at = seq(from = 0, to = 21, by = 7), cex = 1.2, col = "grey80") #Add x-axis
mtext(side = 1, "% DEG", line = 3, cex = 1.5) #Add x-axis label

#dev.off()
```















