---
title: "WGCNA"
author: "Yaamini Venkataraman"
date: "7/8/2020"
output: github_document
---

In this script, I will conduct a WGCNA to understand *Z. marina* gene expression patterns. The code is based on [ tutorials from the WGCNA package](https://horvath.genetics.ucla.edu/html/CoexpressionNetwork/Rpackages/WGCNA/Tutorials/index.html).

# Set up R Markdown file

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Install packages

```{r}
#install.packages("dichromat")
install.packages("BiocManager")
BiocManager::install("WGCNA")
```

```{r}
require(dichromat)
require(WGCNA)
options(stringsAsFactors = FALSE) #Important setting to include from tutorial
```

# Obtain session information

```{r}
sessionInfo() #Obtain session information
```

# Format data

Relevant tutorial: https://horvath.genetics.ucla.edu/html/CoexpressionNetwork/Rpackages/WGCNA/Tutorials/FemaleLiver-01-dataInput.pdf

## Import log CPM data

```{r}
zosteraAllGenes <- read.delim("../EdgeR/EXP.CON.Zostera.logCPM.txt", header = TRUE, quote = "") #Import data with log CPM information for all expressed genes
head(zosteraAllGenes) #Confirm import
```

```{r}
zosteraExpr0 <- as.data.frame(t(zosteraAllGenes[,-1])) #Transpose dataframe to format for WGCNA package. Do not include the first column of gene IDs
names(zosteraExpr0) <- zosteraAllGenes$GeneID #Use gene ID as column names
head(zosteraExpr0) #Confirm formatting
```

## Check for missing values and outliers

```{r}
gsg <- goodSamplesGenes(zosteraExpr0, verbose = 3) #Check for genes and samples with too many missing values
gsg$allOK #TRUE = all genes have passed test, can continue
```
 
```{r}
sampleTree <- hclust(dist(zosteraExpr0), method = "average") #Create dendogram to visually inspect for outliers
plot(sampleTree) #No obvious outliers, can proceed
```

## Create clinical trait data matrix

```{r}
zosteraTraits <- data.frame("exposed" = c(rep(1, times = 7), rep(0, times = 8)),
                            "control" = c(rep(0, times = 8), rep(1, times = 7))) #Create a dataframe with clinical trait (laby exposed vs. not exposed) information
rownames(zosteraTraits) <- rownames(zosteraExpr0) #Use same rownames as expression data to create an analagous matrix
head(zosteraTraits) #Confirm dataframe creation
```

```{r}
sampleTree2 <- hclust(dist(zosteraExpr0), method = "average") #Create a dendogram to look at sample and trait clustering
traitColors <- numbers2colors(zosteraTraits, signed = FALSE) #Convert traits to color values
plotDendroAndColors(sampleTree2, traitColors,
                    groupLabels = names(zosteraTraits)) #Plot dendogram. S_2A funky
```

```{r}
save(zosteraExpr0, zosteraTraits, file = "zostera-WGCNA-dataInput.RData") #Save expression and trait data
```

# Network connection and module detection

Relevant tutorial: https://horvath.genetics.ucla.edu/html/CoexpressionNetwork/Rpackages/WGCNA/Tutorials/FemaleLiver-02-networkConstr-man.pdf

## Determine soft-thresholding power

```{r}
#lnames <- load(file = "zostera-WGCNA-dataInput.RData") #Load data input from previous section, if needed
```


```{r}
powers <- c(c(1:10), seq(from = 12, to = 20, by = 2)) #Create a set of soft-thresholding powers
sft <- pickSoftThreshold(zosteraExpr0, powerVector = powers, verbose = 5) #Use the network topology analysis function to evaluate soft-thresolding power values
```

```{r}
plot(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
     xlab = "Soft Threshold (power)", ylab = "Scale Free Topology Model Fit,signed R^2", type = "n",
     main = paste("Scale independence")) #Plot scale-free topology fit index as a function of the soft-thresholding power
text(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2], labels = powers, cex = 1, col = "red") #Add sft values
abline(h=0.90,col="red") #Add line corresponding to a R^2 cutoff of h = 0.90. Lowest power that reaches 0.90 is sft = 14
```

```{r}
plot(sft$fitIndices[,1], sft$fitIndices[,5],
     xlab = "Soft Threshold (power)", ylab = "Mean Connectivity", type = "n",
     main = paste("Mean connectivity")) #Plot mean connectivity as a function of soft-thresholding power
text(sft$fitIndices[,1], sft$fitIndices[,5], labels = powers, cex = 1, col = "red") #Add sft values. 14 still looks good
```

## Co-expression similarity and adjacency

```{r}
softPower <- 14 #Set soft power based on plots above
adjacency <- adjacency(zosteraExpr0, power = softPower) #Calculate adjacencies using softPower = 14
```

## Topological Overlap Matrix (TOM)

```{r}
TOM <- TOMsimilarity(adjacency) #Minimize noise and spurious associations by transforming adjacency into TOM
dissTOM <- 1-TOM #Calculate dissimilarity matrix
```

## Clustering using TOM

```{r}
geneTree <- hclust(as.dist(dissTOM), method = "average") #Create hierarchical clustering object
plot(geneTree, xlab = "", sub = "", main = "Gene clustering on TOM-based dissimilarity",
     labels = FALSE, hang = 0.04) #Plot initial dendogram. Dissimilarity is based on topological overlap
```

```{r}
minModuleSize <- 30 #Set the minimum module size (# of genes that need to be in a module). Default from WGCNA tutorial
dynamicMods <- cutreeDynamic(dendro = geneTree, distM = dissTOM,
                             deepSplit = 2, pamRespectsDendro = FALSE,
                             minClusterSize = minModuleSize) #Cut the branches of the dendogram to identify WGCNA modules
table(dynamicMods) #Look at the table of modules. There are 18 modules, and 272 genes that do not fit into a module (label 0). Modules have between 42 and 557 genes
```

```{r}
dynamicColors <- labels2colors(dynamicMods) #Convert module numbers into colors
table(dynamicColors) #Look at the list of colors
```

```{r}
plotDendroAndColors(geneTree, dynamicColors, "Dynamic Tree Cut",
                    dendroLabels = FALSE, hang = 0.03,
                    addGuide = TRUE, guideHang = 0.05,
                    main = "Gene dendrogram and module colors") #Plot dendogram with module colors 
```

## Merge modules with similar expression profiles

Merging allows us to combine modules with genes that are highly co-expressed. To do this, we can create eigengenes and cluster them.

```{r}
MEList = moduleEigengenes(zosteraExpr0, colors = dynamicColors) #Calculate eigengenes
MEs <- MEList$eigengenes #Save eigengenes as a new object
MEDiss <- 1-cor(MEs) #Calculate dissimilarity of eigengenes
METree <- hclust(as.dist(MEDiss), method = "average") #Create cluster object
```

```{r}
plot(METree, main = "Clustering of module eigengenes",
     xlab = "", sub = "") #Plot dendogram of clustered eigengenes
dynamicMergeCut(15) #Identify cut height based on sample number (15)
MEDissThres <- dynamicMergeCut(15) #Set cut height threshold based on the number of samples. 1-MEDissThres = correlation
abline(h = MEDissThres, col = "red") #Add cut line
```

```{r}
merge <- mergeCloseModules(zosteraExpr0, dynamicColors, cutHeight = MEDissThres, verbose = 3) #Use an automatic merging function to merge eigengenes
mergedColors <- merge$colors #Extract merged module colors
mergedMEs <- merge$newMEs #Extract eigengenes of merged modules
```

```{r}
plotDendroAndColors(geneTree, cbind(dynamicColors, mergedColors),
                    c("Dynamic Tree Cut", "Merged dynamic"), 
                    dendroLabels = FALSE, hang = 0.03, 
                    addGuide = TRUE, guideHang = 0.05) #Plot the dendogram with original and merged eigengene colors
```

## Rename and save variables for subsequent analysis

```{r}
moduleColors <- mergedColors #Rename to moduleColors
colorOrder = c("grey", standardColors(50)) #Determine color order
moduleLabels = match(moduleColors, colorOrder)-1 #Construct numerical lables based on colors
MEs <- mergedMEs #Replace unmerged MEs
save(MEs, moduleLabels, moduleColors, geneTree, file = "zostera-WGCNA-networkConstruction.RData")
```

# Relate modules to external clinical traits (*L. zosterae* exposed vs. control)

Relevant tutorial: https://horvath.genetics.ucla.edu/html/CoexpressionNetwork/Rpackages/WGCNA/Tutorials/FemaleLiver-03-relateModsToExt.pdf

## Quantify module-trait associations




















